Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-15 paper/paper.tex
typos...

We  need the  following proposition that  connects  liveness analysis  with
minefield semantics transitions. 
\begin{proposition}
\label{prop:correctness}
\begin{enumerate} 
\item Consider a \LET\ definition \linebreak $:\pi: (\LET~x
  \leftarrow~\ldots~\IN~e$, such that there is a use of $x$
  in $e$, i.e.  an occurrence of $\LET~y~\leftarrow~s~in
  \ldots$, where $s$ is $(\CAR\ x)$, $(\CDR\ x)$ or $(+\ x\ y)$, or
  $(\SIF\ x\ e1\ e2)$ or $(\SRETURN\ x)$. Then the demand on $x$ at $\pi$ is non-null. 
\item Assume that a  \LET\ definition $\pi: (\LET~z  \leftarrow~(\CONS~x~y)~\IN~e$ has an occurrence of 
 $(\CAR~z)$ in $e$. Then the liveness of $x$ at $\pi$ includes $\epsilon$.  Similarly if $e$ contains $(CDR~z)$, then the liveness of$y$ at $\pi$ includes $\epsilon$.  
\end{enumerate},
\end{proposition}

@@

\begin{proof}
For a program  executing according to {\em minefield  semantics}, if a
\bang\ has to  occur then some variable which is  dereferenced must be
set to  $\bot$, between its  creation and  use points.  This  can only
happen if  the liveness of  the variable being dereferenced  was null.
From  the lemma~\ref{lemma:no-calls-correctness}  it  is clear  that this  cannot
happen.
\end{proof}

@@

\newcommand{\hgt}{3cm}
\begin{figure*}[p]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
   & {\epsfig{file=lambda_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=treejoin_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory usage.  Column (a) shows complete  usage while column
   (b) enlarges  a part  of the  usage.  The blue  and the  red curves
   indicate the number of cons cells  in the active semi-space for RGC
   and LGC  respectively.  The  black curve  represents the  number of
   reachable cells and  the grey curve represents the  number of cells
   that are  actually live (of  which liveness analysis does  a static
   approximation).   x-axis   is  the  time  measured   in  number  of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by factor $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

Memory   usage   graphs   for   select   benchmarks   in   are   shown
Figure~\ref{fig:memory-usage}, Column  (a).  As the number  of garbage
collections tend to  be very large, for each benchmark  we also show a
window which  is representative  of the  behavior for  that particular
benchmark (Column (b)).  In all the programs we can see that the curve
corresponding to  LGC (red  line) regularly dips  below the  RGC curve
(blue line), the dip is specially visible for the benchmarks where LGC
outperforms  RGC substantially.   The  graphs also  include curve  for
reachable cells (black,  obtained approximately by forcing  RGC to run
at a very high frequency) and the live cells for particular run (grey,
obtained by post processing the heap access at the end of a program).

>>>>

We  need the  following proposition that  connects  liveness analysis  with
minefield semantics transitions. 
\begin{proposition}
\label{prop:correctness}
\begin{enumerate} 
\item   Consider  a   \LET\  definition:   \linebreak  $\pi\!\!\!:   (\LET~x
  \leftarrow~\ldots~\IN~e$, such  that there is  a use of $x$  in $e$,
  i.e.  an occurrence of $\LET~y~\leftarrow~s~in \ldots$, where $s$ is
  $(\CAR\ x)$, $(\CDR\ x)$  or $(+\ x\ y)$, or $(\SIF\  x\ e1\ e2)$ or
  $(\SRETURN\ x)$. Then the demand on $x$ at $\pi$ is non-null.
\item    Assume    that    a   \LET\    definition    $\pi:    (\LET~z
  \leftarrow~(\CONS~x~y)~\IN~e$  has an  occurrence  of $(\CAR~z)$  in
  $e$.  Then  the  liveness  of  $x$  at  $\pi$  includes  $\epsilon$.
  Similarly  if $e$  contains $(CDR~z)$,  then the  liveness of$y$  at
  $\pi$ includes $\epsilon$.
\end{enumerate}
\end{proposition}

@@

\begin{proof}
For a program  executing according to {\em minefield  semantics}, if a
\bang\ has to  occur then some variable which is  dereferenced must be
set to  $\bot$, between its  creation and  use points.  This  can only
happen if  the liveness of  the variable being dereferenced  was null.
From the lemma~\ref{lemma:no-calls-correctness} it  is clear that this
cannot happen. {\color{Myblue}  Too bad - self  reference!! Should this
  be Proposition~\ref{prop:correctness}? }
\end{proof}

@@

\newcommand{\hgt}{2.9cm}
\begin{figure*}[p]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
   & {\epsfig{file=lambda_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=treejoin_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory usage.  Column (a) shows complete  usage, column
   (b) enlarges  a part  of the  usage.  The blue  and the  red curves
   indicate the number of cons cells  in the active semi-space for RGC
   and LGC  respectively.  The  black curve  represents the  number of
   reachable cells and  the lightblue curve represents the  number of cells
   that are  actually live (of  which liveness analysis does  a static
   approximation).   x-axis   is  the  time  measured   in  number  of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by  $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

Memory   usage   graphs   for   select   benchmarks   in   are   shown
Figure~\ref{fig:memory-usage}, Column  (a).  As the number  of garbage
collections tend to  be very large, for each benchmark  we also show a
window which  is representative  of the  behavior for  that particular
benchmark (Column (b)).  In all the programs we can see that the curve
corresponding to  LGC (red  line) regularly dips  below the  RGC curve
(blue line), the dip is specially visible for the benchmarks where LGC
outperforms  RGC substantially.   The  graphs also  include curve  for
reachable cells (black,  obtained approximately by forcing  RGC to run
at a very high frequency) and the live cells for particular run (lightblue,
obtained by post processing the heap access at the end of a program).
